Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-09 paper/related.tex
Edits.

An ownership type system for safe region-based memory management in
real-time Java has been proposed by~\cite{MIT03}. Like us, they too
assume a source language with programmer-managed memory regions, and
focus on proving safety of programs written in that language. Their
source language admits various kinds of regions in order to support
shared-memory concurrency. However, all their regions have
lexically-scoped lifetimes. In contrast, we admit regions with
dynamically determined lifetimes in order to support message-passing
concurrency. We borrow outlives relation from their formal
development, and our type system bears some similarities to their's.
However, our language also admits parametric polymorphism (generics)
and higher-order functions, whose interaction is non-trivial in
context of a region type system. On the other hand, our type system
does not admit region kinds and ownership, and is considerably simpler
than theirs. Moreover, our type system comes with a type
safety result that formalizes its guarantees with respect to a
well-defined operational semantics. Like Cyclone, ~\cite{MIT03}'s
language is explicitly typed, with some support for local type
inference. Consequently, region types for methods and classes need to
be written explicitly. In contrast, we support full type inference
that eliminates such need.

~\cite{HMN01} proposes a flow-sensitive region-based memory management
for first-order programs that proposes to overcome some of the
drawbacks of~\cite{tofte97} by generalizing~\cite{tofte97}'s approach
to regions with dynamic lifetimes. However, dynamic regions are still
not first-class values of the language. Nonetheless, due to aliasing
among region names, ~\cite{HMN01} requires reference counting for
regions to ensure memory safety.

>>>>

An ownership type system for safe region-based memory management in
real-time Java has been proposed by~\cite{MIT03}. Like us, they too
assume a source language with programmer-managed memory regions, and
focus on proving safety of programs written in that language. Their
source language admits various kinds of regions in order to support
shared-memory concurrency. However, all their regions have
lexically-scoped lifetimes. In contrast, we admit regions with
dynamically determined lifetimes in order to support message-passing
concurrency. We borrow outlives relation from their formal
development, and our type system bears some similarities to their's.
However, our language also admits parametric polymorphism (generics)
and higher-order functions, whose interaction is non-trivial in
context of a region type system. On the other hand, our type system
does not admit region kinds and ownership, and is considerably simpler
than theirs. Furthermore, our type system comes with a type
safety result that formalizes its guarantees with respect to a
well-defined operational semantics. Like Cyclone, ~\cite{MIT03}'s
language is explicitly typed. Although there is some support for local
type inference, region types for methods and classes need to be
written explicitly. In contrast, we support full type inference that
eliminates any such need.

~\cite{HMN01} proposes a flow-sensitive region-based memory management
for first-order programs that proposes to overcome some of the
drawbacks of~\cite{tofte97} by generalizing~\cite{tofte97}'s approach
to regions with dynamic lifetimes. However, dynamic regions are still
not first-class values of the language, and reference counting is
nonetheless needed to ensure memory safety.~\cite{WW01} extends lambda
calculus with first-class regions with dynamic lifetimes, and imposes
restrictive linear typing to control accesses to regions. All type
annotations need to be written explicitly. Moreover, invariants needed
to ensure safety under region transfers, such as the absence of
references that escape a transferable region, cannot be enforced
under~\cite{WW01}'s type system.
